\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2016-07-20 redaction/dhp_draft_july2016.tex
Rédaction dans le github

\date{\today}
\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX. Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01. Finally, any remaining errors are ours.}}}
%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa), Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine (DIAL - UMR 225 \& LEDa)} } \maketitle

\bigskip

\begin{abstract}

\begin{spacing}{1.2}

{\normalsize XXXX}

{\normalsize \bigskip }

\end{spacing}
\end{abstract}

@@

\noindent Trade costs remain central in international economic analysis. Defined as the costs associated with the exchange of goods across national borders, they are a major obstacle to international economic integration. In addition, they might be the key to the explanation of major puzzles in international macroeconomics (\citealp{rogoff}). Historical perspective show that indeed, trade costs variations are a major determinant of the level of trade flows. According to the estimates by \citet{Jacks08}, trade cost declines explain around 55 percent of the pre World War I trade boom and 33 percent of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter war trade collapse. After 1950, average trade costs fell by 16\%. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000.

Trade costs are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff-costs), time costs (time to ship goods) and transport costs \emph{per se}. Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \citet{Glaeser04} find that, over the twentieth century, the costs of moving goods have declined by over 90\% in real terms. However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%. Studies overviewed by \citet{Behar_Venables} are consistent with those results: the fall in measured transport costs has been relatively small. The importance of fuel costs and technical progress (which improved speed and reliability rather than decreased costs) in transport explain that result.

Despite the decrease in distance-related costs\footnote{\citet{Behar_Venables} suggest freight costs account for two thirds of the effect of distance, the remaining third ebmoyding more or less the effect of time}, several papers (mostly based on empirical estimates of the gravity equation) in the last decade have shown that distance remained a major obstacle to trade (e.g. \citealp{Head_Mayer04} and \citealp{Disdier_Head08}). \citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs (1.7 $=$ 1.55 $\times$ 1.74 - 1). Again, the latter comprises 44\% border-related trade barriers and 21\% transportation costs. That means that transportation costs are higher than each component of border costs taken separately (namely, 8\% for tariff and non-tariff costs, 7\% for language differences, 14\% for for currency differences and 9\% for other costs).\footnote{1.44 $=$ 1.08 $\times$ 1.07 $\times$ 1.14 $\times$  1.09)}. Therefore, the share of transport costs in the consumer price of manufactured goods remains high (\citealp{Lafourcade_Thisse}). Besides, \citet{combes_lafourcade} emphasize that the figures provided by \citet{anderson_wincoop_jel} should be considered as a a downward benchmark, and could be higher in developing economies. Despite technical progress and infrastructure development, transport costs remain large and deserve attention. In that respect, \citet{Behar_Venables} support that the elasticity of trade with respect to freight costs is around -3, which remains a large number.

Following \citet{samuelson1954}, standard models of international trade have usually relied on the iceberg cost hypothesis, implicitly reducing trade costs to their transport part, and modelling the latter as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded). This implies that more expensive goods are costlier to trade. However, many common empirical facts support the existence of additive costs: pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure (see \citealp{Irrazabal_2015}). The additive versus iceberg trade costs is not a new debate in the international trade literature. The famous Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: the demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out''). \citet{martin2012} gives a strong empirical support to this conjecture: based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite. \citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: the elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption. Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs. Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level. Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%. The additive structure of trade costs emerges only when weight is forced to be independent of price. Besides, \citet{Lashka15} supports that higher-priced, still exported to faraway markets, varieties are heavier and involve proportionally higher shipping costs, the latter increasing the relative demand for high-markup varieties. Markups would therefore be the driving force for explaining the ``shipping the good apples out'' effect in the presence of \emph{ad-valorem} trade costs.

A correct measure of transport costs is also important for understanding several key features of international trade. For example, the existence of additive costs may explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows. Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e. more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments. Besides, several recent papers show that the existence of additive trade costs has important non trivial effects when assessing the welfare effects of trade liberalization. \citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive (per-unit) trade costs, in addition to iceberg/multiplicative costs. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs. The difference arises from dissimilar adjustments at the extensive margin - fewer firms enter into the export market following a reduction in additive costs. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff.

Our purpose in this paper is to provide results on the size of transportation costs over time, explicitly distinguishing between multiplicative and additive parts. Our paper stands at the crossroad of these two strands of the literature, the first one dealing with the dynamics of transport, and more generally trade costs over time, the second one focusing on the measurement and nature of the latter. More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over a large time dimension - several decades. To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}. Therefore, our period of study runs from 1974 to 2013. Closely related to this paper is the work by \citet{Irrazabal_2015}, who develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level data, their results suggest that additive costs are on average 14 percent, relatively to the median price, and that they are strongly correlated with standard proxies for trade costs (like e. g., distance). However, while our data requires that we focus only on transportation costs, our approach departs from theirs in several key aspects.

First, our theoretically agnostic approach aims at letting the data speak, and provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models; to our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the shares of multiplicative and additive costs in total transport costs. Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs. Third, the time dimension of our data allows us to characterize the evolution of transportation costs, and more precisely the dynamics of the relative share of each specific (multiplicative and additive) cost over a forty years time span; to our knowledge, no firm-level database is able to provide such a time coverage. Fourth, we are also able to discriminate our results by transport mode (air and vessel).

Finally, our analysis is also indirectly related to a large strand of the international trade literature assessing the impact of trade costs on international exchanges. Following \citet{anderson_wincoop} and \citet{anderson_wincoop_jel}, there has been a huge bunch of articles relying on gravity equations to infer indirectly trade costs as tariff equivalents, or losses related to various kinds of trade barriers - many of these works are surveyed in \citet{anderson_wincoop}; among the various subsequent works, see in particular \citet{anderson_yotov}, \citet{novy13} or \citealp{chen_novy}). However, our purpose in the present paper is different. We first infer directly transport costs from price data, discriminating between multiplicative and additive parts. In a second step, following an approach similar to \citet{Behar_Venables}, we relate the previous measures to traditional determinants of transport costs (like e. g., distance), in order to understand how differently the latter impact the multiplicative and additive transport costs.

%time as a trade barrier, on risque d'avoir des problèmes avec ça... \citet{Lafourcade_Thisse},

XXX Here a summary of our results and contribution XXX

The next section presents the data and our fundamental equation. Section 3 reports our baseline estimates for multiplicative and additive measures of transport costs across time, as well as a number of robustness checks. Section 4 discusses the results of estimates relating traditional determinants of transport costs to the previous measures, while Section 5 XXX has yet to be discussed!!! XXX. The last section concludes.

%\noindent More generally,  Among the closest to our own work, Eaton et al. (2002) develop a Ricardian general equilibrium trade model to estimate parameters relating to absolute advantage, to comparative advantage, and to geographic barriers. Anderson and Wincoop (2003) elaborate a theory-founded gravity equation to assess the cost of trade frictions, and use it to provide a direct measure of the trade losses due to national borders (between 20 and 50 percent).

\section{Data Sources and Empirical Methodology}

@@

\noindent

Our strategy for representing transportation costs is relatively standard, and consists in exploiting the difference between commodity-level export and prices. We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, origins and transportation mode. More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs. The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price. Strictly higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.

To have the import price of the good (CIF price), next to the export price (FOB price) is highly valuable. Indeed, this allows us thus to estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs. This differentiates us from \citet{Irrazabal_2015}, that can only estimate the ratio of additive costs as a share of the tola price. However, note that this measure based on \emph{freight} cost ignores another dimension of transportation costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit.

\subsection{Database}

%laius sur pourquoi pas de donnes de firmes (triop comutational), ni comtrade (trop mauvais, citer RIE) XXX

\noindent Even if the data of transportation costs we rely could be available at the firm-level, the use of a nonlinear estimator (see below) triggers computational limitations that do not make them a likely option. A product-level bilateral database of CIF-FOB ratios is theoretically possible, and could be built from the well-known UN's COMTRADE data. Unfortunately, \citet{hummels_rie} show that issues inherent to the construction of data (especially the use of ``mirror'' flows) make this database unreliable for our purpose.

We decide therefore to use database for a single country, but with strong statistical reliability arising from a single, trustworthy customs origin. More precisely, the empirical analysis relies on yearly data from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013. Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system. In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air). The first will be crucial to compute transport costs (see below), the second will alow us enlightening potential different dynamics of transport costs across transportation mode.

Due to the abovementioned computational limitations, data are reaggregated at the 3- and 4-digit level: comparing different levels of aggregation will be useful to check differences and the presence of biases precisely due to aggregation. Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products.

% Why the gap CIF-FOB is a good measure of trade costs (cf mimeo by Gaulier and Zignago)? JH: to me, This is not the right question to ask. We have the freight costs in our data. The true question is: how well these freight costs proxy correctly transportation costs as a whole.

\subsection{Key empirical specification}

\noindent Our purpose is to provide estimates over time of the relative shares of multiplicative and additive costs among total transport costs. To do so, we start from a very simple equation, quite alike the one proposed by \citet{hummels2010}:

\begin{equation}\label{eq:base}
p = \tau \widetilde{p} + t
\end{equation}

or as a ratio

\begin{equation}\label{eq:base_bis}
\frac{p}{\widetilde{p}} = \tau  + \frac{t}{\widetilde{p}}
\end{equation}

This equation relates the ratio of the consumer price (ie, cif price, denoted $p$) to the producer price (ie, fob price, denoted $\widetilde{p}$) to the two types of costs (multiplicative and additive). Iceberg trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs) while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs).
%\begin{equation*}
%p =\tau \widetilde{p}+t
%\end{equation*}
We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis. Let us denote $i$ the origin country and $k$, the sector. Hence, the equation to be estimated is given by:

\begin{equation}\label{eq:base_estimee}
\frac{p_{ik}}{\widetilde{p}_{ik}} = \tau_{ik}+\frac{t_{ik}}{ \widetilde{p}_{ik}}
\end{equation}

\section{Decomposing the Dynamics of Transport Costs over Time}

\subsection{Estimation Strategy}

% nonlinear least squares: problem with heteroskedasticity???

We transpose the approach by \citet{Irrazabal_2015} by considering that both multiplicative and additive costs are separable between origin country and products, in a multiplicative way for the former and an additive way for the latter. In other words, $\tau$ and $t$ from equation~\ref{eq:base_estimee} become:

%\begin{eqnarray}
%\tau_{ik} &=& dist^{\rho_I}F^I_i \label{eq:iceberg}\\
%t_{ik} &=& dist^{\rho_A}F^A_i-1 \label{eq:add}
%\end{eqnarray}

@@

%\begin{equation}
%\frac{p_{ik}}{\widetilde{p}_{ik}} =dist^{\rho_I}F^I_ik+\frac{dist^{\rho_A}F^A_ik-1}{ \widetilde{p}_{k}}\label{eq:estimatedequation}
%\end{equation}

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} =\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$. Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term. More precisely:

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right)\times \varepsilon_{ik}
\end{equation}

Log-linearization gets the following, tractable empirical equation:

\begin{equation}
ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= ln(\varepsilon_{ik})$ .

>>>>

\title{\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}} %\thanks{We are especially grateful to XXX. Part of this research was funded by the French Agence Nationale de la Recherche (ANR), under grant ANR-11-JSH1 002 01. Finally, any remaining errors are ours.}}}
%We also thank participants at the EEA 27\textsuperscript{th} Congress (Oslo), ETSG 13\textsuperscript{th} meetings (Copenhagen), INFER Workshop (Orleans), 45\textsuperscript{th} Congress of the Canadian Economic Association (Ottawa), IAW Workshop (Tubingen), and seminars at WTO, Strathclyde University, OFCE and EQUIPPE-University of Lille.
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris Dauphine, LEDa \& PSL, Chercheur associ\'{e} - Sciences Po (OFCE)}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS (UMR 9221), Universit\'{e} de Bretagne Occidentale-ICI (EA 2652) and CEPII}\qquad Lise \textsc{Patureau}\thanks{Universit\'{e} Paris Dauphine, LEDa \& PSL} }
 \maketitle

\date{July 2016}
\bigskip

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent. However, many common empirical facts support the existence of additive costs. This paper provides results on the size and composition of transport costs over time, explicitly distinguishing between its multiplicative and additive components, using sectorial cif-fob prices over 1974-2013 taken from US imports data. First, we estimate the two components of transport costs, by transport mode (air or ocean). We thus find that ocean transport costs decompose in an ad-valorem tax equivalent of mean value 3.22\% and a per-unit cost that amounts to 2.85\% of the mean fob price over the period. These values are respectively equal to 2.5\% and 1.9\% in air shipping. Second, we provide a quantitative assessment what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2. On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%. Third, the time dimension of our data allows us to characterize the evolution of transport costs. After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012. While this order of magnitude applies for both the additive and multiplicative components of vessel transport costs, air transport costs exhibit more contrasted trends.
\end{abstract}

@@

\noindent Trade costs remain central in international economic analysis. Defined as the costs associated with the exchange of goods across national borders, they are usually split into transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. Many believe that the latter have dramatically decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \citet{Glaeser04} find that, over the twentieth century, the costs of moving goods have declined by over 90\% in real terms. However, \citet{hummels2007} shows that the bulk of price declines in transportation comes from air shipping, where average cost per ton-kilometer shipped dropped by 92\% between 1955 and 2004; concerning ocean shipping, which represents the major part of world trade value, decline in trade prices are much less obvious, even if the rise in containerization lowered shipping costs from 3 to 13\%. Studies overviewed by \citet{Behar_Venables} are consistent with those results: The fall in measured transport costs has been relatively small. According to these authors, this is attributable to the growing importance of fuel costs and technical progress, which improved speed and reliability rather than decreased costs.\smallskip

Trade costs are considered as a major obstacle to international economic integration and international trade flows. According to the estimates by \citet{Jacks08}, trade cost declines explain around 55\% of the pre-World War I trade boom and 33\% of the post-World War II trade boom, while the abrupt rise in trade costs explains the inter-war trade collapse. After 1950, average trade costs fell by 16\%, notably through the reduction of policy trade costs promoted through the GATT (WTO starting in 1995) multilateral agreements. Based on panel data, \citet{novy13} thus finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Yet, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs (typically captured through distance) still remain a major obstacle to trade (e.g. \citealp{Head_Mayer04} and \citealp{Disdier_Head08}). \citet{anderson_wincoop_jel} estimate that average trade costs for industrialized countries put up to 170\% markups over production costs, divided into 55\% distribution costs and 74\% international costs. Within this latter dimension, 44\% are border-related trade barriers and 21\% transportation costs. \citet{Lafourcade_Thisse} also find that the share of transport costs in the consumer price of manufactured goods remains high, while \citet{Behar_Venables} obtain that the elasticity of trade with respect to freight costs is sizeable, by around -3. If much of trade-policy barriers have been removed over the second half of the 20th century, these findings suggest that the transport costs component of the overall trade costs remain large and deserve attention. This is accordingly the focus of the paper.\smallskip

Following \citet{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an \emph{ad valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, the ``iceberg cost'' hypothesis). However, many common empirical facts support the existence of additive costs. As documented by \citet{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure. The structure (additive vs multiplicative) of transport costs is far from being anecdotal, as the literature has long pointed out its role in shaping the pattern of trade flows. The Alchian and Allen conjecture (\citealp{alchian}), which points out that the relative price of two varieties of some good will depend on the level of trade costs, does rely on the existence of additive costs: The relative demand for more expensive/higher quality product goods should increase with trade cost (``shipping the good apples out''). \citet{martin2012} gives a strong empirical support to this conjecture: Based on a very disaggregated firm-product-level database of French exporters, he finds that firms charge higher fob unit values on exports to more remote countries, whereas the iceberg hypothesis would imply the opposite. \citet{hummels_skiba} also find some strong evidence in favor of the Alchian-Allen conjecture: The elasticity of freight rates with respect to price is estimated to be well below unity, in contradiction with the iceberg assumption. Also, their estimates implied that doubling freight costs increases average fob export prices by 80-141 percent, consistent with high quality goods being sold in markets with high freight costs. \textbf{je ne sais pas quoi faire de la reference a lashkaripour. Complique d'en parler, car on ne peut pas le contredire car on n'a pas les donnees qui le permettent. Ne pas en parler? C'est un peu limite.} The existence of additive costs may also explain a large number of zeros in bilateral trade flows, and more generally, the granularity of trade flows. Relying on Spanish and US transaction-level trade data, \citet{Hornok14} find that additive trade costs of are associated with less frequent and larger shipments, i.e. more ``lumpiness'', in international trade; in other words, exporters wait to fill completely a container before sending it abroad, to decrease as much as possible the number of shipments.

%Conversely, \citet{Lashka15} finds strong empirical support for the iceberg hypothesis when considering the data at a sufficiently disaggregated level. Based on US imports data, he finds that a 10\% increase in the price of an item increases the shipping cost by 9.5\%. The additive structure of trade costs emerges only when weight is forced to be independent of price. Besides, \citet{Lashka15} supports that higher-priced, still exported to faraway markets, varieties are heavier and involve proportionally higher shipping costs, the latter increasing the relative demand for high-markup varieties. Markups would therefore be the driving force for explaining the ``shipping the good apples out'' effect in the presence of \emph{ad-valorem} trade costs.

Beyond the positive aspect of understanding trade patterns, several recent papers also point out the normative implications of additive trade costs. \citealp{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the iceberg component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in multiplicative costs. Calibrating on Norwegian firm-level data for 2004, \citet{Irrazabal_2015} find that an additive import tariff reduces welfare and trade by more than an identically-sized multiplicative tariff. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much can be done in quantifying such gains though, by lack of an empirical characterization of the additive component of trade costs.\footnote{The one exception being \citet{Irrazabal_2015}, upon which we come back later.} One objective of the paper is to palliate this gap. \bigskip

Our paper contributes to the literature by providing results on the size of transport costs over time, explicitly distinguishing between multiplicative and additive parts. More precisely, we provide estimates of the relative importance of additive and multiplicative transport costs over several decades. To do so, we update the detailed US customs, sector-level data from the US Imports of Merchandise used by \citet{hummels2007}, to cover the period from 1974 to 2013. Closely related to this paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level data, their results suggest that additive costs are on average 30-45\% relative to the median consumer price, and that they are strongly correlated with standard proxies for trade costs (like e. g., distance). However, while our data requires that we only consider transportation costs, our approach departs from theirs in several key aspects.

First, our theoretically agnostic approach provides a fairly simple framework for assessing \emph{both} multiplicative and additive parts of transportation costs, which may prove to represent a non-negligible advantage for calibrating related models. We thus obtain that the mean values over 1974-2012 of iceberg costs are equal to 2.5\% in air and 3.2\% in ocean shipping, whereas the additive component amounts to 1.9\% and 2.85\% of the fob price in air and ocean shipping respectively. To our best knowledge, our paper is the first to provide such an extensive quantitative assessment of the magnitude of both multiplicative and additive costs in total transport costs. Second, we provide, through standard measures of ``goodness-of-fit'', an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2. On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%. Third, the time dimension of our data allows us to characterize the evolution of transport costs, by transport mode (air or vessel) and for each specific (multiplicative and additive) cost over a forty years time span. After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012. Decomposing by transport mode and by type of transport costs allows to go deeper into this picture. First, as in \cite{hummels2007}, we find that the transport costs decrease is more pronounced in air than in ocean shipping, starting in 1984. However, while ocean shipping costs display a regular decreasing pattern until 2012, air transport costs rather get stabilized over 2005-2012. Second, while both the additive and multiplicative components of ocean transport costs have roughly decreased by 40\% over the period, air transport costs exhibit more contrasted trends. If the reduction of air shipping costs can most be attributed to the decrease in its multiplicative component over 1974-2005, the opposite scheme prevails afterwards, the reduction in per-unit air transport costs being compensated by an increase in the ad-valorem component. \smallskip

Section \ref{sec:data_method} explains the data sources and the empirical methodology retained in the paper. Sections \ref{sec:results_decomposition} and \ref{sec:results_trends} report our results. In Section \ref{sec:results_decomposition}, we characterize the role of the additive component of transport costs. After reporting the mean values over the period (by transport mode), we show the improved performance in including additive trade costs in the measure of transport costs. This being established, Section \ref{sec:results_trends} characterizes the trends in each component of transport costs (by transport mode) over the period 1974-2012. Section \ref{sec:conclu} concludes.

\section{Data Sources and Empirical Methodology \label{sec:data_method}}

@@

As in \cite{hummels2007} (among others), our measure of transportation costs consists in exploiting the difference between commodity-level export and import prices.
We first use values, quantities and freight costs to recover free-on-board (FOB) and cost-insurance-fret (CIF) prices, by goods, country of origin and transportation mode. More precisely, the (unit) FOB price is computed as the total customs value divided by the shipping weight; in other words, it is the price for the good net of transportation costs. The CIF price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the CIF price divided by the FOB price. Strictly higher than 1, the variable provides therefore with a measure of transport costs as a proportion of the good's price, an \emph{ad valorem} equivalent.

The database we use to construct our measure of transport comes from US annual Imports of Merchandise provided by the Census bureau\footnote{More information available at: \url{http://www.census.gov/foreign-trade/reference/products/catalog/fl_imp.txt}}, spanning from 1974 to 2013. Using this dataset has (at least) three main advantages. First, this dataset has been used by \cite{hummels2007}, which enables us to compare our results to his findings. However, we complement Hummels' (2007) findings as we extend the time coverage to 2013 (while \cite{hummels2007} stops in 2004). As we show in Section \ref{sec:results_trends}, extending the time period over he recent years delivers interesting insights regarding the trends in air shipping costs. Second, and importantly, this dataset delivers a strong statistical reliability arising from a single, trustworthy customs origin. Based on customs declarations, this dataset inventories all imports (both values and quantities) by origin to the United states at the HS 10-digit highly disaggregated level, with a concordance code to the SITC 5-digit coding system. In addition, the database reports information regarding freight expenditures and transportation mode (Ocean Vessel and Air). The first will be crucial to compute transport costs (see below), the second will allow us enlightening substantial differences in the dynamics of transport costs across transportation mode. Third, using this dataset allows us to have the import price of the good (CIF price), next to the export price (FOB price). This is highly valuable, as we can estimate both the \textit{levels} of the iceberg trade costs and of the additive trade costs. This differentiates us from \citet{Irrazabal_2015}, which can only estimate the ratio of additive costs as a share of the total consumer price (the only price they can encover).

It is also true that using this dataset has drawbacks. First, our measure of the cif-fob price gap only covers transportation costs by nature, thereby being mute about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}.  Further, in terms of transport costs \textit{per se}, this measure being based on freight costs, omits the other dimension of transport costs related to the time value of goods in transit. According to \citet{anderson_wincoop_jel}), the 21\% markup over production costs coming from transport costs includes both directly measured freight costs and 9\% tax equivalent of the time value of goods in transit. \textbf{en faire un contre-argument positif}

Even if the data of transportation costs is available at a more disaggregated level, we use sectorial price data at the 3-digit level, primarily for technical reasons. As detailed below, the use of a nonlinear estimator triggers computational limitations that do not make them a likely option, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level (for some selected years). Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. Depending on the considered year, this leaves us with around 200 (3-digits) and 600-700 (4-digits) products, from approximately 200 countries of origin.

\subsection{Empirical specification}

\paragraph{The estimated equation} Our purpose is to provide estimates over time of the sizes of multiplicative and additive costs among total transport costs. To do so, we start from a very simple equation (similar to \cite{Irrazabal_2015} or \cite{martin2012}, among others):

\begin{equation}
p = \tau \widetilde{p} + t \label{eq:base}
\end{equation}

\noindent This equation expressed the consumer (import, or cif) price $p$ as a function of the producer (export, or fob) price $\widetilde{p}$ given both per-unit ($t$) and ad-valorem ($\tau$) transport costs. As usual in the literature, the so-called ``iceberg'' trade costs are denoted $\tau$ (with $\tau=1$ meaning no iceberg trade costs), while additive trade costs are labeled $t$ (with $t=0$ implying no additive trade costs). We estimate this equation for each year over the period 1974-2013, and for each of the two transportation modes reported (air or vessel), on a sectoral-origin country basis. Let us denote $i$ the origin country, and $k$, the sector (or product). Transforming the above equation (\ref{eq:base}) as ratio, we thus get the equation to be estimated as given by:\footnote{Also keeping in mind that the estimated equation is transport-mode and year specific (air or ocean shipping). We skip the year and transport-mode dimensions here to alleviate notations.}

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation}

\paragraph{Estimation Strategy} We follow \citet{Irrazabal_2015} by considering that \textit{1°)} both multiplicative and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{2))} in a multiplicative way for the former and an additive way for the latter. In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) become:

@@

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a ``one-lower bond'', since by construction, the \emph{cif} price $p$ cannot be lower than the \emph{fob} price: $p>\widetilde{p}_{ik}$. Taking into account this constraint in the estimation requires to impose a multiplicative structure for the error term, according to:

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \varepsilon_{ik}
\end{equation*}

Taking in log, this finally drives us to estimate the following equation:

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= \ln(\varepsilon_{ik})$.
